\documentclass[12pt]{article}

% --- Packages ---
\usepackage[T1]{fontenc}        % Better font encoding
\usepackage{mathpazo}           % Palatino font for text and math
\usepackage{amsmath, amssymb}   % Math packages
\usepackage{graphicx}           % For including graphics
\usepackage{setspace}           % For line spacing
\usepackage{geometry}           % For page margins
\usepackage{verbatim}           % For displaying code or console output
\usepackage{parskip}            % Better paragraph spacing

% --- Formatting ---
\geometry{margin=1in}
\setstretch{1.15}

\begin{document}

\title{PhD Industrial Organization: Homework 2}
\author{Franco Diaz Laura}
\date{\today}
\maketitle

\section{Question 1}

\textbf{How to run code:}\\
To run the code, navigate to the \texttt{src} directory and execute:
\begin{verbatim}
julia Rust_replication.jl
julia run_results.jl
\end{verbatim}


The following is the output (note that it takes approximately \textbf{2248 seconds} to run):

\begin{quote}\ttfamily
=== Rust (1987) — NFXP Step 2 (fast) ===

$\beta = 0.9999$, $\theta_{30} = 0.3919$, $\theta_{31} = 0.5953$ ($p_2 = 0.0128$)\\
Bins: 90, $\Delta = 5000$ miles\\
Estimates: RC = 10.0428, $\theta_{11} = 0.456457$\\
Negative log-likelihood (nll): 163.584577
\end{quote}

So my estimates for RC are close but $\theta_{11}$ is not that close.

\section{Question 2}
\paragraph{(b) CCP (Hotz--Miller) two--step estimator for the Rust (1987) bus problem.}
Let $x\in\mathcal{X}=\{1,\ldots,90\}$ index mileage bins (5k miles), and $i\in\{0,1\}$ index choices
($i=0$ keep, $i=1$ replace). The discount factor $\beta$ and transition matrices
$P(y\mid x,i)$ are taken as given from Step~1. Idiosyncratic shocks are i.i.d.\ type~I extreme value.
We seek $\theta=(RC,\theta_{11})$ in the linear flow cost $c(x;\theta_{11})=0.001\,\theta_{11}\,x_{\text{miles}}$ so that
\[
u(x,i;\theta)=
\begin{cases}
-RC - c(0;\theta_{11})=-RC, & i=1 \ (\text{replace}),\\[2pt]
-c(x;\theta_{11}), & i=0 \ (\text{keep}).
\end{cases}
\]

\noindent\textbf{Step 1 (nonparametric CCPs).}
For each state $x$, compute smoothed empirical conditional choice probabilities (CCPs)
\[
\widehat p(i\mid x)=\frac{n_{xi}+\alpha}{n_x+J\alpha},\qquad J=2,\ \alpha>0,
\]
where $n_x$ is the number of observations with state $x$ and $n_{xi}$ the number of times choice $i$ is taken at $x$
(Laplace smoothing avoids $\log 0$ when actions are rare).

\noindent\textbf{Step 2 (inclusive value from CCPs).}
With logit errors and the baseline $i_0=0$ (keep) normalized, the ex--ante value satisfies
$EV(x)=\gamma+\log\sum_j \exp v_j(x)$ and, using $\sum_j \exp v_j(x)=1/\widehat p(i_0\mid x)$,
\[
\widetilde{EV}(x)\equiv -\log\widehat p(i_0\mid x),
\]
which equals $EV(x)$ up to the constant $\gamma$; the constant cancels in differences below.

\noindent\textbf{Step 3 (Hotz--Miller inversion).}
For each $x$ and each $i\neq i_0$ form
\begin{align*}
\Delta_{\text{HM}}(x,i)
&= \underbrace{\log\widehat p(i\mid x)-\log\widehat p(i_0\mid x)}_{\text{current-period log-odds}}
\;-\;
\beta\Big(
\underbrace{\sum_{y}\widetilde{EV}(y)\,P(y\mid x,i)}_{\text{next-period ex-ante value under }i}
-
\underbrace{\sum_{y}\widetilde{EV}(y)\,P(y\mid x,i_0)}_{\text{under baseline }i_0}
\Big).
\end{align*}
For the Rust utility, the structural difference is
\[
u(x,i;\theta)-u(x,i_0;\theta)=
\begin{cases}
-RC + 0.001\,\theta_{11}\,x_{\text{miles}}, & i=1\ \text{vs }i_0=0.
\end{cases}
\]

\noindent\textbf{Step 4 (minimum distance / NLS estimation).}
Choose $\theta$ to minimize
\[
Q(\theta)=\sum_{x\in\mathcal{X}}\sum_{i\neq i_0} w_{xi}\,
\big[\Delta_{\text{HM}}(x,i) - \{u(x,i;\theta)-u(x,i_0;\theta)\}\big]^2,
\]
with weights $w_{xi}$ (e.g.\ $w_{xi}=n_x/\sum_x n_x$). If $u$ is linear in $\theta$, this is OLS;
otherwise solve by nonlinear least squares (or GMM).

\medskip
\noindent\textbf{ NPL refinement (Aguirregabiria--Mira).}
Initialize $p^{(0)}(i\mid x)=\widehat p(i\mid x)$. For $k=0,1,2,\ldots$ repeat:
\begin{enumerate}\setlength{\itemsep}{2pt}
\item $\widetilde{EV}^{(k)}(x)=-\log p^{(k)}(i_0\mid x)$.
\item Given $\theta^{(k)}$, form model CCPs
\[
p^{\text{model}}(i\mid x;\theta^{(k)}) \;\propto\; \exp\!\left(u(x,i;\theta^{(k)})+\beta\sum_y \widetilde{EV}^{(k)}(y)P(y\mid x,i)\right).
\]
\item Update CCPs with damping: $p^{(k+1)}:=\rho\,p^{(k)}+(1-\rho)\,p^{\text{model}}(\cdot\mid\cdot;\theta^{(k)})$.
\item Re-estimate $\theta^{(k+1)}$ by the criterion $Q(\theta)$ using $p^{(k+1)}$.
\item Stop when $\max_{x,i}\lvert p^{(k+1)}(i\mid x)-p^{(k)}(i\mid x)\rvert<\text{tol}$.
\end{enumerate}

\end{document}
